### 2012-12-17 annualreports/2012/balancesheet.tex
2012 annual report

\section{Balance Sheet}
31\superscript{st} March 2012

\begin{center}
\begin{tabular}{  l  r  r r r}
\toprule
                            & \multicolumn{2}{c}{2012} &  \multicolumn{2}{c}{2011} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
                                               & £        & £       & £      & £      \\
{\bf Fixed Assets}                             &          &         &        &        \\
{\bf Tangible fixed assets}                    &          & 5,149   &        & 5,525  \\
\cmidrule(r){2-3} \cmidrule(r){4-5} \\
{\bf Current Assets}                           &          &         &        &        \\
Rental Deposit                                 & 7,696    &         & 7,696  &        \\
Inventory                                      & 163      &         & 0      &        \\
Cash at bank and in hand                       & 14,889   &         & 7,688  &        \\
Creditors: amounts falling due within one year & -167     &         & -2,511 &       \\
Prepaid Subscriptions                          & 0        &         & -610   &       \\
{\bf Net current assets}                       &          & 27,898  &        & 12,263 \\
\cmidrule(r){2-3} \cmidrule(r){4-5} \\
{\bf Total assets less current liabilities}    &          & 27,732  &        & 17,787 \\
\cmidrule(r){2-3} \cmidrule(r){4-5}  \\
{\bf Capital and Reserves}&\\
Income and expenditure account                 &          & 27,732  &        & 17,787 \\
\bottomrule
\end{tabular}

\end{center}

{\footnotesize For the financial period ended 31 March 2011 the
company was entitled to exemption from audit under section 477 of the
Companies Act 2006. No member of the company has deposited a
notice, pursuant to section 476, requiring an audit of these
financial statements under the requirements of the Companies Act
2006.

The directors acknowledge their responsibilities for ensuring that
the company keeps accounting records which comply with section 386
of the Act and for preparing financial statements which give a true
and fair view of the state of affairs of the company as at the end
of the financial period and of its profit or loss for the financial
period in accordance with the requirements of sections 394 and 395
and which otherwise comply with requirements of the Companies Act
2006 relating to accounts, so far as applicable to the company.

These financial statements have been prepared in accordance with
the special provisions relating to small companies within Part 15
of the Companies Act 2006 and with the Financial Reporting Standard
for Smaller Entities (effective April 2008).}

\vfill

Approved by the Board for issue on TBC
\vspace{48pt}

Russell Garrett \\
{\bf Director} \\
Company Number: 06807563